NLTK: The Natural Language Toolkit

### 2002-03-20 technical/acl2002/submission.tex
- Rewrote modules section as plaintext - Minor changes - Fixed bugs in preformatted code (needed backslashes, etc)
svn/trunk@513

@@

\textbf{Ease of Use.} The primary purpose of the toolkit is to allow
students to concentrate on building NLP systems.  The more time
students must spend learning to use the toolkit, the less useful it
is.

\textbf{Consistency.} The toolkit should use consistent data structures
and interfaces.

\textbf{Extensibility.} The toolkit should easily accommodate new
components, whether those components replicate or extend the toolkit's
existing functionality.  Thus, the toolkit's design should be modular,
with simple and well-defined interfaces between modules.  The toolkit
should also be structured in such a way that it is obvious where new
extensions would fit into the toolkit's infrastructure.

\textbf{Documentation.} The toolkit, its data structures, and its
implementation all need to be documented.  Three documentation types
are necessary: technical reports, to explain the toolkit's
design and implementation; tutorials, to teach students how to use the
toolkit to perform specific tasks; and reference documentation, to describe
every module, interface, class and method.  All nomenclature must be
carefully chosen and consistently used.

@@

\textbf{Simplicity.} The toolkit should structure the complexities of
building NLP systems, not hide them.  Therefore, each class defined by 
the toolkit should be simple enough that a student could implement it
by the time they finish the course.

\textbf{Modularity.} The interaction between different components of the
toolkit should be kept to a minimum.  In particular, it should be
possible to use complete individual projects using small parts of the
toolkit, without worrying about how they interact with the rest of the
toolkit.  This will allow students to learn to use the toolkit
incrementally throughout the course.  Modularity also makes it easier
to change or extend the toolkit.

@@

\textbf{Comprehensiveness.} The toolkit is not intended to provide a
comprehensive set of tools.  Indeed, there should be a wide variety of 
ways in which students can extend the toolkit.

\textbf{Efficiency.} The toolkit does not need to be highly optimized for
runtime performance.  However, it should be efficient enough that
students can use their NLP systems to perform real tasks.

\textbf{Cleverness.} Clear designs and implementations are far preferable 
to ingenious yet indecipherable ones.

@@

\textbf{Interfaces.}
Python does not directly implement interfaces.  The toolkit
therefore implements interfaces as simple classes, all of whose methods 
raise \texttt{AssertionError} exceptions.  A description of the
functionality provided by the interface is contained in the
class's documentation string.  Each method's documentation string
specifies the behavior of that method.
In order to implement an interface, a class should include that
interface as one of its bases, and should override every
(non-optional) method defined by the interface.
Classes that define interfaces are named with a trailing ``I,'' such
as \texttt{TokenizerI} or \texttt{EventI}.

\textbf{Typing.}
Type checking is an important safety device, allowing programmers to
quickly locate errors in their code.  This is especially important for
novice programmers, who have less experience with debugging.  Every
function and method defined by the toolkit will check the types of its
its arguments.  Given the impact on performance, the level of type checking
will be adjustable.

\textbf{Advanced Language Features.}
The use of advanced language features and programming styles could
potentially make the toolkit much more difficult to use or to
understand.  The fewer language features that a student must learn in
order to use the toolkit, the faster they can start developing NLP
systems.  The language features used by
the toolkit are summarized by figure \ref{fig:feature}.

\begin{figure}
\begin{centering}
\begin{tabular}{|l|l|}
\hline
\textbf{Language Feature} & \textbf{Use} \\
\hline
Classes & used \\
Exceptions & error conditions only \\
Operator Overloading & used \\
Python 2.0 Features & used \\
Python 2.1 Features & not used \\
Function Arguments & minimal use \\
Lambda Functions & minimal use \\
Mapping and Filtering & not used \\
List Comprehensions & used \\
Tkinter & used \\
\hline
\end{tabular}\\
\end{centering}
\caption{Advanced language features used by NLTK.}
\label{fig:feature}
\end{figure}

\section{Modules}

NLTK currently includes the following modules.

\subsection*{Basics}

\begin{description}
\item[token:] Basic classes for processing individual
      elements of text, such as words or sentences. 
\item[tree:] Classes for representing tree structures
      over text (such as syntax trees and morphological trees).
\item[probability:] Classes that encode frequency
      distributions and probability distributions.
\end{description}

\subsection*{Tagging}

\begin{description}
\item[tagger:] A standard interface to tag each token of
      a text with supplementary information, such as its part of
      speech; and several implementations of that interface.
\end{description}

\subsection*{Parsing}

\begin{description}
\item[parser:] A standard interface to produce trees
      representing the structure of texts.
\item[chartparser:] A flexible parser implementation
      that uses a \emph{chart} to record hypotheses about
      syntactic constituents.
\item[srparser\_template:] A partial implementation of a
      shift-reduce parser; completing the implementation was a
      student exercise.
\item[chunkparser:] A standard interface for robust
      parsers used to identify non-overlapping linguistic groups
      (such as noun phrases) in unrestricted text.
\item[rechunkparser:] A regular-expression based
      implementation of the chunk parser interface.
\end{description}

@@

\begin{description}
\item[classifier:] A standard interface for classifying
      texts into categories.
\item[classifier.feature:] A standard way of encoding
      the information used to make classification decisions.
\item[classifier.naivebayes:] A text classifier
      implementation based on the Naive Bayes assumption.
\item[classifier.maxent:] An implementation of the
      maximum entropy model for text classification; and
      implementations of the GIS and IIS algorithms for training
      the classifier.
\item[classifier.featureselection:] A standard interface
      for choosing which features are relevant for making
      classification decisions.
\end{description}

\begin{description}
\item[draw.tree:] A graphical representation for tree
      structures, such as syntax trees and morphological trees. 
\item[draw.tree\_edit:] A graphical interface used to
      build and modify tree structures.
\item[draw.plot\_graph:] A graphical tool to graph
      arbitrary functions.
\item[draw.chart:] An interactive graphical tool used to
      experiment with chart parsers.
\end{description}

@@

\begin{sv}
cascade = [
  ChunkRule('<DT><NN.*><VB.><NN.*>'),
  ChunkRule('<DT><VB.><NN.*>'),
  ChunkRule('<.*>'),
  UnChunkRule('<IN|VB.*|CC|MD|RB.*>'),
  UnChunkRule("<,|\.|``|''>"),
  MergeRule('<NN.*|DT|JJ.*|CD>', '<NN.*|DT|JJ.*|CD>'),
  SplitRule('<NN.*>', '<DT|JJ>')
]
\end{sv}

@@

\begin{sv}
cascade = [
  ChunkRule('<\$|CD|DT|EX|PDT
             |PRP.*|WP.*|\#|FW
             |JJ.*|NN.*|POS|RBS|WDT>*')
]
\end{sv}

@@

The main program which surrounds the student's work is moderately
simple as well.  The line replaced by students is underlined.

\begin{sv}
from nltk.token import *
from nltk.chunkparser import *
from nltk.rechunkparser import *
import os

@@

sent_tok = RETokenizer('(=======+)?\\n\\n+', 0)
sentences = []
files = os.listdir(path)
for file in files:
    print 'Loading %s...' % file
    text = open(path+file).read()
    sentences += sent_tok.tokenize(text, source=file, unit='s')

>>>>

% Note: the maximum paper length is 6 pages.

% Outline:
%   - abstract
%   - introduction
%   - choice of programming language
%   - design criteria
%   - python implementation (??)
%   - modules
%   - uses
%     - assignments/projects
%     - class demonstrations
%   - evaluation
%   - other approaches
%   - conclusion

% To do:
%   - FSA module description
%   - visualization modules description
%   - in-class demonstrations section
%   - Get rid of Python implementation section, or at least shorten it
%   - General clean-up

@@

\noindent\textbf{Ease of Use.} The primary purpose of the toolkit is
to allow students to concentrate on building NLP systems.  The more
time students must spend learning to use the toolkit, the less useful
it is.

\noindent\textbf{Consistency.} The toolkit should use consistent data
structures and interfaces.

\noindent\textbf{Extensibility.} The toolkit should easily
accommodate new components, whether those components replicate or
extend the toolkit's existing functionality.  Thus, the toolkit's
design should be modular, with simple and well-defined interfaces
between modules.  The toolkit should also be structured in such a way
that it is obvious where new extensions would fit into the toolkit's
infrastructure.

\noindent\textbf{Documentation.} The toolkit, its data structures,
and its implementation all need to be documented.  Three documentation
types are necessary: technical reports, to explain the toolkit's
design and implementation; tutorials, to teach students how to use the
toolkit to perform specific tasks; and reference documentation, to
describe every module, interface, class and method.  All nomenclature
must be carefully chosen and consistently used.

@@

\noindent\textbf{Simplicity.} The toolkit should structure the
complexities of building NLP systems, not hide them.  Therefore, each
class defined by the toolkit should be simple enough that a student
could implement it by the time they finish the course.

\noindent\textbf{Modularity.} The interaction between different
components of the toolkit should be kept to a minimum.  In particular,
it should be possible to complete individual projects using small
parts of the toolkit, without worrying about how they interact with
the rest of the toolkit.  This will allow students to learn to use the
toolkit incrementally throughout the course.  Modularity also makes it
easier to change or extend the toolkit.

@@

\noindent\textbf{Comprehensiveness.} The toolkit is not intended to
provide a comprehensive set of tools.  Indeed, there should be a wide
variety of ways in which students can extend the toolkit.

\noindent\textbf{Efficiency.} The toolkit does not need to be highly
optimized for runtime performance.  However, it should be efficient
enough that students can use their NLP systems to perform real tasks.

\noindent\textbf{Cleverness.} Clear designs and implementations are
far preferable to ingenious yet indecipherable ones.

@@

% [EL] This is the only section that I think we might want to keep
% from python impl.  But I'm not sure we even need to keep this one.
\noindent\textbf{Interfaces.}
Python does not directly implement interfaces.  The toolkit
therefore implements interfaces as simple classes, all of whose methods 
raise \texttt{AssertionError} exceptions.  A description of the
functionality provided by the interface is contained in the
class's documentation string.  Each method's documentation string
specifies the behavior of that method.
In order to implement an interface, a class should include that
interface as one of its bases, and should override every
(non-optional) method defined by the interface.
Classes that define interfaces are named with a trailing ``I,'' such
as \texttt{TokenizerI} or \texttt{EventI}.

% [EL] Mention that Python doesn't provide native type checking?  Or
% ditch this section? :)
\noindent\textbf{Typing.}
Type checking is an important safety device, allowing programmers to
quickly locate errors in their code.  This is especially important for
novice programmers, who have less experience with debugging.  Every
function and method defined by the toolkit will check the types of its
its arguments.  Given the impact on performance, the level of type checking
will be adjustable.

%\noindent\textbf{Advanced Language Features.}
%The use of advanced language features and programming styles could
%potentially make the toolkit much more difficult to use or to
%understand.  The fewer language features that a student must learn in
%order to use the toolkit, the faster they can start developing NLP
%systems.  The language features used by
%the toolkit are summarized by figure \ref{fig:feature}.

%\begin{figure}
%\begin{centering}
%\begin{tabular}{|l|l|}
%\hline
%\noindent\textbf{Language Feature} & \textbf{Use} \\
%\hline
%Classes & used \\
%Exceptions & error conditions only \\
%Operator Overloading & used \\
%Python 2.0 Features & used \\
%Python 2.1 Features & not used \\
%Function Arguments & minimal use \\
%Lambda Functions & minimal use \\
%Mapping and Filtering & not used \\
%List Comprehensions & used \\
%Tkinter & used \\
%\hline
%\end{tabular}\\
%\end{centering}
%\caption{Advanced language features used by NLTK.}
%\label{fig:feature}
%\end{figure}

\section{Modules}
% [EL]: I think that this section works better as a textual
% description, rather than a long list.  Also, we're going to be short
% on space.. :)
% What order should these be in?  Put more impressive stuff at the
% beginning and end?  Maybe
% parsing;tagging;visualization;classification?  or swap viz and
% classification?

The toolkit is implemented as a collection of independant
\emph{modules}, each of which defines a specific data structure or
task.

A set of core modules defines basic data types and processing tasks
that are used througout the toolkit.  The \texttt{token} module
defines basic classes for processing individual elements of text, such
as words or sentences.  The \texttt{tree} module provides data
structures for representing tree strucutres over text, such as syntax
trees and morphological trees.  The \texttt{probability} module
defines classes that encode frequency distributions and probability
distributions, including a variety of smoothing techniques.

The remaining modules define data structures and interfaces for
performing specific NLP tasks.

\subsection*{Parsing Modules}

The \texttt{parser} module defines a high-level interface for
producing trees that represent the structure of texts.  The
\texttt{chunkparser} module defines a subinterface for parsers that
identify non-overlapping linguistic groups (such as noun phrases) in
unrestricted text.

Three modules provide implementations for these abstract interfaces.
The \texttt{srparser} module implements a simple shift-reduce parser.
The \texttt{chartparser} module provides a flexible parser that uses a
\emph{chart} to record hypotheses about syntactic constituents.  And
the \texttt{rechunkparser} module defines a transformational
regular-expression based implementation of the chunk parser interface.

\subsection*{Tagging Modules}

The \texttt{tagger} module defines a standard interface for augmenting
each token of a text with supplementary information, such as its part
of speech or its WordNet synset tag; and provides several
different implementations for that interface.

\subsection*{Finite State Automata}

\texttt{fsa}.  Planning to add HMMs??

\texttt{draw.tree}
\texttt{draw.tree\_edit}
\texttt{draw.plot\_graph}
\texttt{draw.chart}
\texttt{draw.fsa}

@@

The \texttt{classifier} module defines a standard interface for
classifying texts into categories.  This interface is currently
implemented by two modules: the \texttt{classifier.naivebayes} module
defines a text classifier based on the Naive Bayes assumption; and the
\texttt{classifier.maxent} module defines the maximum entropy model
for text classification, and includes two algorithms for training the
model: Generalized Iterative Scaling and Improved Iterative Scaling.

The \texttt{classifier.feature} module provides a standard encoding
for the information that is used to make decisions for a particular
classification task.  This standard encoding allows students to
experiment with the differences between different text classification
algorithms, using identical feature sets.

The \texttt{classifier.featureselection} module defines a standard
interface for choosing which features are relevant for a particular
classification task.  Good feature selection can signifigantly improve
classification performance.

%NLTK currently includes the following modules.

%\subsection*{Basics}

%\begin{description}
%\item[token:] Basic classes for processing individual
%      elements of text, such as words or sentences. 
%\item[tree:] Classes for representing tree structures
%      over text (such as syntax trees and morphological trees).
%\item[probability:] Classes that encode frequency
%      distributions and probability distributions.
%\end{description}

%\subsection*{Tagging}

%\begin{description}
%\item[tagger:] A standard interface to tag each token of
%      a text with supplementary information, such as its part of
%      speech; and several implementations of that interface.
%\end{description}

%\subsection*{Parsing}

%\begin{description}
%\item[parser:] A standard interface to produce trees
%      representing the structure of texts.
%\item[chartparser:] A flexible parser implementation
%      that uses a \emph{chart} to record hypotheses about
%      syntactic constituents.
%\item[srparser\_template:] A partial implementation of a
%      shift-reduce parser; completing the implementation was a
%      student exercise.
%\item[chunkparser:] A standard interface for robust
%      parsers used to identify non-overlapping linguistic groups
%      (such as noun phrases) in unrestricted text.
%\item[rechunkparser:] A regular-expression based
%      implementation of the chunk parser interface.
%\end{description}

%\subsection*{Text Classification}

%\begin{description}
%\item[classifier:] A standard interface for classifying
%      texts into categories.
%\item[classifier.feature:] A standard way of encoding
%      the information used to make classification decisions.
%\item[classifier.naivebayes:] A text classifier
%      implementation based on the Naive Bayes assumption.
%\item[classifier.maxent:] An implementation of the
%      maximum entropy model for text classification; and
%      implementations of the GIS and IIS algorithms for training
%      the classifier.
%\item[classifier.featureselection:] A standard interface
%      for choosing which features are relevant for making
%      classification decisions.
%\end{description}

%\subsection*{Visualization}

%\begin{description}
%\item[draw.tree:] A graphical representation for tree
%      structures, such as syntax trees and morphological trees. 
%\item[draw.tree\_edit:] A graphical interface used to
%      build and modify tree structures.
%\item[draw.plot\_graph:] A graphical tool to graph
%      arbitrary functions.
%\item[draw.chart:] An interactive graphical tool used to
%      experiment with chart parsers.
%\end{description}

@@

\begin{sv}
cascade = [
  ChunkRule('<DT><NN.*><VB.><NN.*>'),
  ChunkRule('<DT><VB.><NN.*>'),
  ChunkRule('<.*>'),
  UnChunkRule('<IN|VB.*|CC|MD|RB.*>'),
  UnChunkRule("<,|{\textbackslash}.|``|''>"),
  MergeRule('<NN.*|DT|JJ.*|CD>', '<NN.*|DT|JJ.*|CD>'),
  SplitRule('<NN.*>', '<DT|JJ>')
]
\end{sv}

@@

\begin{sv}
cascade = [
  ChunkRule('<{\textbackslash}\$|CD|DT|EX|PDT
             |PRP.*|WP.*|{\textbackslash}\#|FW
             |JJ.*|NN.*|POS|RBS|WDT>*')
]
\end{sv}

@@

% [EL] I'm not sure this is necessary:
The main program which surrounds the student's work is moderately
simple as well.  The line replaced by students is underlined.
% [EL] I find the undeline hard to read; change to italics or bold or
% vbar in the margin?  Or maybe a \Rightarrow of some sort in the margin?
\begin{sv}
from nltk.token import *
from nltk.chunkparser import *
from nltk.rechunkparser import *
import os

@@

sent_tok = RETokenizer('(=======+)?{\textbackslash}{\textbackslash}n{\textbackslash}{\textbackslash}n+', 0)
sentences = []
files = os.listdir(path)
for file in files:
    print 'Loading \%s...' \% file
    text = open(path+file).read()
    sentences += sent_tok.tokenize(text, source=file, unit='s')
